El Mayor Postseismic \author{Trever Hines} \rule{\headwidth}{1.0pt}

### 2016-03-24 RoughDraft/ElMayor.tex
new commit

\usepackage{amssymb,amsmath}
\usepackage[margin=1.0in]{geometry}
\usepackage{fancyhdr} % required for custom header
\usepackage{graphicx}
\usepackage{listings}
\usepackage{courier}
\usepackage[usenames,dvipsnames]{color}

@@

\begin{document}
\maketitle
\section*{Introduction}

@@

Stations which recorded signals that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, which record deformation that is largely anthropogenic. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake. While several stations were installed after the EMC earthquake to improve the spatial resolution of postseismic deformation \cite{Spinler2015}, our inverse method uses postseismic displacements rather than velocities (e.g. ...), which requires the knowledge of the stations preseismic position. Despite our inability to utilize potentially rich data, we prefer to use postseismic displacements rather than potentially dubious estimates of postseismic velocities.

@@

It is important to note that the shown uncertainties in $u(t)$ do not account for our uncertainty in eq. \ref{TimeSeriesModel}.  For example, we assume that the background rate of deformation can be approximated as having a constant rate, while mechanical models would suggest this to not be true on earthquake cycle timescales \cite{Thatcher1983}. Although based on inspection,  assuming a constant rate of background deformation appears to be appropriate for all stations except for perhaps the stations closest to the Hector Mine epicenter, where postseismic deformation persists.  Also, our model for seasonal deformation in eq. \ref{TimeSeriesModel} assumes a constant amplitude over time, which means that years of particularly heavy or light rainfall cannot be adequately described.  This deficiency in our seasonal model causes our estimate of $u(t)$ to describe some of the unmodeled oscillations (figure \ref{P619PS}).

@@

In this paper, we seek to find the mechanisms driving five years of postseismic deformation following the El Mayor-Cucapah earthquake. We consider afterslip and viscoelastic relaxation in the lithosphere and/or asthenosphere as candidate mechanisms.  Poroelastic rebound has also been used to model postseismic deformation (cite); however, \cite{Gonzalez-ortega2014} suggest that any contribution to postseismic deformation from poroelastic rebound would be negligable.  Additionally, we are considering stations which are sufficiently far away from the main rupture that poroelastic rebound should be insignificant.

In this study we estimate coseismic and time dependent postseismic fault slip, both of which are assumed to occur on the fault geometry estimated by \cite{Wei2011a} with some modifications.  Field studies \cite{Fletcher2014} and LIDAR observations \cite{Oskin2012} have revealed a significantly more complicated fault geometry than what was inferred by \cite{Wei2011a}, especially within the Sierra Cucapah.  However, we find the relatively simple coseismic fault geometry from \cite{Wei2011a} to be sufficient because most of the stations used in this station are sufficiently far from the El Mayor-Cucapah epicenter that they are insensitive to the details in the fault geometry found by \cite{Fletcher2014} and \cite{Oskin2012}.  We do, however, modify the fault geometry from \cite{Wei2011a} by extending segment 2 northward by 40 km which is motivated by the clustering of aftershocks on the northern tip of coseismic rupture zone \cite{Kroll2013} \cite{Hauksson2011}.  This extended fault segment was also found to be necessary by \cite{Rollins2015} and \cite{Pollitz2012} for describing postseismic deformation.

We consider a variety of rheologic models for the lithosphere and asthenosphere in this study.  The simplest rheologic model would be to consider the lithospere and asthenosphere to be effectively elastic. In such case, the rheologic parameters, consisting of the Lame parameters, $\lambda$ and $\mu$, and density, $\rho$, are reasonably well known, leaving the only unknowns to be the distribution of fault slip, which can be easily estimated due to the linearity of the forward problem. \cite{Rollins2015} found that postseismic deformation following the El Mayor-Cucapah earthquake can be well explained with afterslip on the coseismicly ruptured fault, which has been extended down to the lithosphere-asthenosphere boundary inferred by \cite{Lekic2011}, and they did not require any viscoelastic relaxation to describe the observations.  They did however, find that the amount of afterslip required for the southernmost fault segment to be unrealistically large, equivalent to a Mw7.2 earthquake.  Since most of the GPS data is located north of the rupture zone, slip on the Indiviso fault segment is least constrained and could be acting as a proxy for distributed relaxation in the lower crust or upper mantle.

It is well understood that afterslip at sufficiently great depths can explain postseismic deformation resulting from viscoelastic relaxation \cite{Savage1990}, and so in the interest of preventing a non-unique solution, we assuming afterslip to occur only at seismogenic depths (<15 km), and allow for viscoelastic relaxation in the lower crust (15-30 km) depth and the upper mantle.We choose such shallow constraints on afterslip because rate-state friction based models would suggest afterslip to occur at shallow depths to accommodate a coseismic slip deficit \cite{Marone1991}, which is indeed inferred in the coseismic models by \cite{Wei2011a} and (Fialko 2010). Additionally, as noted by \cite{Rollins2015} the postseismic uplift just north of the rupture zone can be well described with slip at seismogenic depths, while afterslip which is constrained to depths beneath 15 km predicts subsidence in that region.

@@

With afterslip constrained to shallow depths, we must now select a viscoelastic rheology for the lower crust and upper mantle. Maxwell viscoelasticity, which can be schematically illustrated as a spring and dashpot connected in series, is the simplest and commonly used rheologic model for studies on postseismic deformation NUR IS SLS! (e.g. \cite{Nur1974}, \cite{Pollitz2000}, \cite{Hetland2003},\cite{Freed2006a}, \cite{Johnson2009}, \cite{Hearn2009}).  Assuming Maxwell viscoelasticity introduces a single the unknown rheologic parameter. A Maxwell viscoelastic model of the lithosphere would predict deformation that decays steadily over time, which fails to describe the multiple timescales of relaxation observed in postseismic deformation \cite{Savage1997}.  An additional deficiency in modeling the lithosphere with a Maxwell rheology is that inferred viscosities tend to depend on the the timescale of the geophysical process being modelled \cite{Peltier1981}. When assuming Maxwell viscoselasticity, upper mantle viscosities for Southern California inferred from postseismic deformation tend to be on the order of $10^{18}$ Pa s (\cite{Pollitz2000}, \cite{Thatcher2008}, \cite{Spinler2015}. However, studies of interseismic deformation \cite{Lundgren2009} and lake loading \cite{Luttrell2007}, which are both processes that occur on the timescale of $~10^2$ years, infer a Maxwell viscoelastic upper mantle with a viscosity on the order of $10^{19}$ Pa s.

To account for the apparently multiple timescales of relaxation, \cite{Pollitz2003} used a Burgers rheology upper mantle to model postseismic deformation following the Hector Mine earthquake and inferred a transient viscosity of $1.6\times10^{17}$ Pa s and a steady state viscosity of $4.6\times10^{18}$ Pa s, which still falls short of the steady state viscosity required for longer term geophysical processes. The particularly weak transient viscosity was found to be necessary to describe the early rapid postseismic deformation, although other researchers have successfully described the early postseismic deformation for the same earthquake with fault creep and poroelastic rebound. Nevertheless, laboratory experiments have demonstrated that mantle rock exhibits two phases of relaxation which can be adequately described with a Burgers rheology \cite{Chopra1997} and we should expect to see some postseismic signal resulting from transience viscoelasticity.

The third linear viscoelastic rheology which we consider is the Zener model, often referred to as a standard linear solid, whose mechanical analogue consists of a spring connected in series with a Kelvin element. A generalization of the Zener model, consisting of a series of Kelvin elements connected in series each of which representing a distinct deformation mechanism, is commonly used to describe the spectrum of frequency dependent attenuation \cite{Liu1976}.  The highest viscosity needed to describe attenuation of earth's lowest frequency normal modes is on the order of $10^{16}$ Pa s \cite{Yuen1982}, which is too weak to be perceptible in the GPS data used in the present study.

As an exploratory step, we also perform a slip inversion, assuming an elastic lithosphere and asthenosphere.  The difference between the analysis in  \cite{Rollins2015} and the present study is that we include a wider range of GPS stations in our inversion and we also perform a temporal estimation of afterslip, as opposed to estimating slip at a single epoch. The material properties used in our inversion are given in table 1 and are consistent with the properties used by \cite{Wei2011a} and the 1D SCEC background velocity model.  We discretize the

 and the constrained and the only para.  where the rheologic parameters       
In order to model postseismic deformation we must ddetermine an appropriate rheologic model for the lithosphere and asthenosphere.  
To model postseismic deformation we must first assume a rheologic model.  The simplest viscoelastic rheologic model would be a Maxwell model, where shear stresses and strains can be described with the mechanical analogy of a spring and dashpot connected in series.  If coseismic and postseismic slip, described by $s(\xi,t)$, occurs on a fault $F$, and viscoelastic relaxation of subsequent stresses occurs in a Maxwell viscoelastic medium $L$, which has a spatially variable viscosity, $\eta(\zeta)$, then coseismic and postseismic deformation can be described by 
\begin{equation}\label{GeneralForward}
  u(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F s(\tau,\xi) f(t-\tau,x,\xi) d\xi d\tau
\end{equation}
where $g(x,\xi)$ is the elastic Green's function relating slip at $\xi$ to instantaneous displacement at $x$ and $f(t,x,\xi)$ describes the rate of deformation at $x$ and time $t$ resulting from viscoelastic relaxation of stresses induced by slip  at $\xi$.  The variables which we are primarily concerned with estimating in this study are $s(\xi,t)$ and $\eta(\zeta)$.  Numerous experimental and geophysical studies suggest that a Maxwell rheology is inappropriate to describe the behaviour of the uppermost mantle \cite{Chopra1997} \cite{Pollitz2003} \cite{Freed2004} and so we do not limit ourselves to exploring Maxwell viscolesticity.  When considering other stress-linear viscoelastic models, such as a Zener or Burgers model, the description of displacements remains unchanged from eq \ref{GeneralForward} except that $f$ is parameterized by additional rheologic parameters which must also be estimated.  The difficulty in  determining an appropriate rheologic model and then estimating $s(\xi,t)$ and the unkown rheologic parameters from observations of $u(x,t)$ is that it is a nonlinear inverse problem with many unknown parameters and a computationally expensive forward problem.  We can greatly simplify the inverse problem by first considering the early postseismic period where we can approximate $u(x,t)$ for any stress-linear viscoelastic rheology as
\begin{equation}\label{ApproxForward}
  u(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F\int_L \frac{s(\tau,\xi)}{\eta_\mathrm{eff}(\zeta)} h(x,\xi,\zeta) d\zeta d\xi d\tau
\end{equation}

>>>>

\usepackage{amssymb,amsmath}
\usepackage[margin=1.0in]{geometry}
\usepackage{fancyhdr} % required for custom header
\usepackage{graphicx}
\usepackage{listings}
\usepackage{courier}
\usepackage[usenames,dvipsnames]{color}
\usepackage[maxfloats=50]{morefloats}

@@

\begin{document}
\maketitle
\section*{Abstract}
we propose a anelastic phase of deformation in the mantle distinct from phases associated with attenuation

\section*{Introduction}

@@

Stations which recorded signals that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, which record deformation that is largely anthropogenic. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake. While several stations were installed after the EMC earthquake to improve the spatial resolution of postseismic deformation \cite{Spinler2015}, our inverse method uses postseismic displacements rather than velocities (e.g. ...)
ELABORATE ON VELOCITY VS DISPLACEMENT
, which requires the knowledge of the stations preseismic position. Despite our inability to utilize potentially rich data, we prefer to use postseismic displacements rather than potentially dubious estimates of postseismic velocities.

@@

It is important to note that the shown uncertainties in $u(t)$ do not account for the non-negligible epistemic uncertainty in eq. \ref{TimeSeriesModel}.  For example, we assume that the background rate of deformation can be approximated as having a constant rate, while mechanical models would suggest this to not be true on earthquake cycle timescales \cite{Thatcher1983}. Although based on inspection,  assuming a constant rate of background deformation appears to be appropriate for all stations except for perhaps the stations closest to the Hector Mine epicenter, where postseismic deformation persists.  Also, our model for seasonal deformation in eq. \ref{TimeSeriesModel} assumes a constant amplitude over time, which means that years of particularly heavy or light rainfall cannot be adequately described.  This deficiency in our seasonal model causes our estimate of $u(t)$ to describe some of the unmodeled oscillations (figure \ref{P619PS}).

@@

In this paper, we seek to find the mechanisms driving five years of postseismic deformation following the El Mayor-Cucapah earthquake. We consider afterslip and viscoelastic relaxation in the lithosphere and/or asthenosphere as candidate mechanisms.  Poroelastic rebound has also been used to model postseismic deformation \cite{Jonsson2003}; however, \cite{Gonzalez-ortega2014} suggest that any contribution to postseismic deformation from poroelastic rebound would be negligable.  Additionally, we are considering stations which are sufficiently far away from the main rupture that poroelastic rebound should be insignificant.

\subsection{Fault Geometry}
In this study we estimate coseismic and time dependent postseismic fault slip, both of which are assumed to occur on the fault geometry estimated by \cite{Wei2011a} with some modifications.  Field studies \cite{Fletcher2014} and LIDAR observations \cite{Oskin2012} have revealed a significantly more complicated fault geometry than what was inferred by \cite{Wei2011a}, especially within the Sierra Cucapah.  However, we find the relatively simple coseismic fault geometry from \cite{Wei2011a} to be sufficient because most of the stations used in this station are sufficiently far from the El Mayor-Cucapah rupture zone that they are insensitive to the details in the fault geometry found by \cite{Fletcher2014} and \cite{Oskin2012}.  We do, however, modify the fault geometry from \cite{Wei2011a} by extending segment 2 northward by 40 km which is motivated by the clustering of aftershocks on the northern tip of coseismic rupture zone \cite{Kroll2013} \cite{Hauksson2011}.  This extended fault segment was also found to be necessary by \cite{Rollins2015} and \cite{Pollitz2012} for describing postseismic deformation.

We consider a variety of rheologic models for the lithosphere and asthenosphere in this study.  The simplest rheologic model would be to consider the lithospere and asthenosphere to be effectively elastic and isotropic. In such case, the rheologic parameters, consisting of the Lame parameters, $\lambda$ and $\mu$, and density, $\rho$, are reasonably well known, leaving the only unknowns to be the distribution of fault slip, which can be easily estimated due to the linearity of the forward problem. \cite{Rollins2015} found that postseismic deformation following the El Mayor-Cucapah earthquake can be well explained with afterslip on the coseismicly ruptured fault, which has been extended down to the lithosphere-asthenosphere boundary inferred by \cite{Lekic2011}, and they did not require any viscoelastic relaxation to describe the observations.  They did however, find that the amount of afterslip required for the southernmost fault segment to be unrealistically large, equivalent to a Mw7.2 earthquake.  Since most of the GPS data is located north of the rupture zone, slip on the Indiviso fault segment is least constrained and could be acting as a proxy for distributed relaxation in the lower crust or upper mantle.

It is well understood that afterslip at sufficiently great depths can explain postseismic deformation resulting from viscoelastic relaxation \cite{Savage1990}, at least in 2D earthquake models. In the interest of preventing a non-unique solution, we assuming afterslip to occur only at seismogenic depths (<15 km), and allow for viscoelastic relaxation in the lower crust (15-30 km) depth and the upper mantle.We choose such shallow constraints on afterslip because rate-state friction based models would suggest afterslip to occur at shallow depths to accommodate a coseismic slip deficit \cite{Marone1991}, which is indeed inferred in the coseismic models by \cite{Wei2011a} and (Fialko 2010). Additionally, as noted by \cite{Rollins2015} the postseismic uplift just north of the rupture zone can be well described with slip at seismogenic depths, while afterslip which is constrained to depths beneath 15 km predicts subsidence in that region.

@@

\subsection{Rheologic Models}
With afterslip constrained to shallow depths, we must now select a viscoelastic rheology for the lower crust and upper mantle. Maxwell viscoelasticity, which can be schematically illustrated as a spring and dashpot connected in series, is the simplest and commonly used rheologic model for studies on postseismic deformation (e.g. \cite{Nur1974}, \cite{Pollitz2000}, \cite{Hetland2003},\cite{Freed2006a}, \cite{Johnson2009}, \cite{Hearn2009}).  The Maxwell model describes both an instantaneously elastic and fluid-like phase of deformation.   Assuming Maxwell viscoelasticity introduces the additional rheologic parameter, $\eta_M$, which is typically estimated for the lower crust and upper mantle as an inverse problem, where the viscosities are chosen to minimize the difference between the observed postseismic deformation and deformation predicted by a mechanical model. A Maxwell viscoelastic model of the lithosphere would predict deformation that decays steadily over time, which fails to describe the observable deformation which exhibits transient deformation which decays rapidly within the first year of an earthquake followed by a relatively steady rate of deformaiton \cite{Savage1997}.  An additional deficiency in describing postseismic deformation with viscoelastic relaxation in a Maxwell viscoselastic lithosphere is that inferred viscosities tend to be considerably lower than viscosities inferred from geophysical processes that occur over longer timescales. For example, postseismic studies in Southern California following the Landers \cite{Pollitz2000}, Hector Mine \cite{Pollitz2001}, and El Mayor-Cucuapah earthquake \cite{Spinler2015}, \cite{Rollins2015}, have all inferred a Maxwell viscoselastic upper mantle with a viscosities ranging from $10^{17}$ to $10^{18}$ Pa s; however, studies on interseismic deformation \cite{Lundgren2009} and lake loading \cite{Luttrell2007}, which are both processes that occur on the timescale of $~10^2$ years, infer a Maxwell viscoelastic upper mantle with a viscosity on the order of $10^{19}$ Pa s.

We also consider the Zener viscoelastic model whose mechanical analogue consists of a spring connected in series with a Kelvin element. The Zener model is used to describe both an instantaneous and delayed phase of elasticity.  A generalization of the Zener model, consisting of numerous Kelvin elements connected in series, is commonly used to describe the spectrum of frequencies where seismic attenuation is observed \cite{Liu1976}.  The highest viscosity needed to describe attenuation of earth's lowest frequency normal modes is on the order of $10^{16}$ Pa s \cite{Yuen1982} and the largest characteristic relaxation time would be on the order of days, which is too short to be perceptible in the GPS data used in the present study. We therefor ignore the delayed elasticity associated with seismic attenuation models, and consider such models to be effectively instantaneously elastic. Nevertheless, we still explore the possibility of describing postseismic deformation with a Zener model that has relaxation timescales of months to years.

Of course, it has long been recognized that a Zener rheology provides an incomplete descriptions of the asthenosphere, as it does not have the fluid behaviour required for mantle dynamics on geologic time scales. \cite{Yuen1982} proposed a Burgers rheology with a low viscosity ($\approx 10^{16}$ Pa s) in the Kelvin element and high viscosity ($\approx 10^{21}$ Pa s) in the Maxwell element which is capable of describing both seismic attenuation and the fluid-like behaviour required for glacial isostatic rebound and mantle convection. The justification of a Burger's rheology mantle is further supported by laboratory experments on the strength of olivine \cite{Chopra1997}. \cite{Pollitz2003} sought to describe postseismic deformation following Hector Mine with a Burgers rheology mantle and they found a best fitting transient viscosity of $1.6\times10^{17}$ Pa s and steady state viscosity of $4.6\times10^{18}$ Pa s. While the Burgers rheology was introduced as a means of bridging to gap between relaxation observed in long and short term geophysical processes, the inferred steady state viscosity from \cite{Pollitz2003} is still inconsistent with the Maxwell viscosities inferred from \cite{Lundgren2009} and \cite{Luttrell2007}. Additionally, the transient viscosity inferred by \cite{Pollitz2003} is needed to describe the earliest phase of postseismic deformation following the Hector Mine earthquake, while \cite{Pollitz2003} ruled out deep afterslip as an alternative mechanism based on inconsistent vertical deformation, it is still possible to successfully describe all components of early postseismic deformation following the Hector Mine earthquake with afterslip at seismogenic depths \cite{Jacobs2002}. In this paper we seek to constrain the rheologic parameters describing a Burgers upper mantle that are consistent with postseismic deformation and long term geophysical processes, while also accounting  for shallow afterslip, which perhaps biased the solution from \cite{Pollitz2003}.

\subsection{Constraint on Effective Viscosity}\label{InitialInversion}
For the three linear viscoelastic rheologies described above, Maxwell, Zener, and Burgers, postseismic displacement resulting from time dependent fault slip, $s(\xi,t)$, on a fault $F$ can be described as  
\begin{equation}\label{GeneralForward}
  u(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F s(\tau,\xi) f(t-\tau,x,\xi) d\xi d\tau
\end{equation}
where $g(x,\xi)$ describes the instantaneous elastic displacements at $x$ resulting from a unit of slip at $\xi$, and $f(t,x,\xi)$ describes the velocity at $x$ resulting from viscoelastic relaxation of stresses induced by slip at $\xi$. $f$ is a function of the properties of the chosen rheologic model, which are generally not well known.  The parameters controlling the instantaneous elastic behaviour of the lithosphere and asthenosphere, $\mu$, $\lambda$, and density $\rho$, are relatively well known and we use the values listed in table 1, which are consistent with the properties used in the coseismic model by \cite{Wei2011a} and the 1D SCEC background velocity model.  The remaining rheologic parameter, which we will assume to have a depth dependence, are estimated to optimize the predicted fit to observable postseismic displacement.

In order greatly simplify the posed inverse problem we use to method describe in \cite{Hines2015} to constrain the effective lithospheric viscosity structure for the early postseismic period.  The method relies upon the fact that immediately after an earthquake, stresses throughout the lithosphere and asthenosphere, denoted at $L$, are controlled by the relatively well known instantaneous elastic properties.  This means that the rate of deformation in each parcel of the lithosphere will deform at a rate that is inversely proportional to its effective viscosity, $\eta_{\mathrm{eff}}=\frac{\sigma}{\dot{\epsilon}}|_{t=0}$, and independent of $\eta_{\mathrm{eff}}$ elsewhere. Then, as a consequently of linear elasticity, the initial rate of surface deformation resulting from viscoelastic relaxation is a linear combination of the surface deformation resulting from each parcel, scaled by the reciprocal of the parcel's viscosity. That is to say   
\begin{equation}\label{InitialRate}
  f(0,x,\xi) = \int_L \frac{h(x,\xi,\zeta)}{\eta_\mathrm{eff}(\zeta)} d\zeta, 
\end{equation}
where $h(x,\xi,\zeta)$ describes the initial rate of deformation resultion from viscoelastic relaxation at $\zeta$ induced by slip at $\xi$ and, most importantly, $h$ is independent of all rheologic properties other than those that control the instantaneous elastic behaviour, $\mu$, $\lambda$ and $\rho$. Subsitution eq. \ref{InitialRate} into \ref{GeneralForward} gives us an approximation for early postseismic deformation,
\begin{equation}\label{ApproxForward}
  u(x,t) \approx \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F\int_L \frac{s(\tau,\xi)}{\eta_\mathrm{eff}(\zeta)} h(x,\xi,\zeta) d\zeta d\xi d\tau,
\end{equation}
which is valid for as long as the rate of deformation resulting from viscoelastic relaxation is approximately constant. Although eq. \ref{ApproxForward} may only valid for a short portion of the postseismic period, its utility becomes apparent when noting that $g$ and $h$ are functions of the fault geometry and instantaneous elastic properties, which are fairly well known, and so $g$ and $h$ can be computed numerically as a preprocessing step and the forward problem, eq. \ref{ApproxForward}, can be rapidly evaluated for any realization of $s$ and $\eta_{\mathrm{eff}}$.  This is in contrast to evaluating the full forward problem, eq. \ref{GeneralForward}, numerically for each realization of $s$ and $\eta_{\mathrm{eff}}$ which is far more computationally demanding.

We perform an initial inversion of postseismic displacements using eq. \ref{ApproxForward} as out forward problem. As noted, we use the fault geometry from \cite{Wei2011}, except that each segment extends to 15 km depth and segment 2 is extended 40 km northward.  Each fault segment is discretized into roughly 4 km by 4 km patches and we estimate a strike-slip and thrust component of slip for each patch. We estimate coseismic slip as well as a rate of afterslip which is constant over the time intervals spanning 0.0-0.125, 0.125-0.25, and 0.5-1.0 years following the earthquake, as well as over 1.0 year intervals for the remaining postseismic period.  We impose that the coseismic slip and rates of afterslip are within 45$^\circ$ of right-lateral slip for all segments except for segment 1, where slip is constrained be within $45^\circ$ of normal.  We estimate $\eta_{\mathrm{eff}}$ within 6 vertically stratified layers which have depths ranging from 0-15 km, 15-30 km, 30-60 km, 60-90 km, 90-120 km, 120-150 km, and 150 km to the bottom of our numerical model domain, 800 km. Details of the inverse method are described in \cite{Hines2015}.  We note that a nonlinear Kalman filter based inverse method can be employed to estimate $s$ and $\eta_{\mathrm{eff}}$ in a manner akin to \cite{Segall1997} or \cite{McGuire2003}, where we do not have to explicitly impose a time dependent parameterization of $s$. Indeed we have thoroughly explored that possibility but we ultimately prefer the method described in \cite{Hines2015} because of its relative simplicity and because we believe the piecewise continuous respresentation of slip with respect to time to be sufficiently general.

The first step in our inverse method is to determine the time during the postseismic period at which point the early postseismic approximation, eq. \ref{ApproxForward}, is no longer approriate, which we will denote as $t_{\mathrm{end}}$.  As noted, it is valid for approximately as long as the rate of deformation resulting for viscoelastic relaxation is approximately constant. We can almost certainly assume that deformation at the most far field stations which are ~400 km away from the El Mayor-Cucapah is the result of viscoelastic relaxation. Then the approximation should be valid for as long as a linear trend adequately approximates the far field deformation. Using this logic, it would appear that $t_{\mathrm{end}}$ is $\sim1$ year after the El Mayor-Cucapah earthquake.  Another way to determine $t_{\mathrm{end}}$ is to find the best fitting prediction of eq. \ref{ApproxForward} to various durations of the postseismic timeseries and determine when eq. \ref{ApproxForward} is not longer capable of describing the deformation without incurring systematic misfits.  This is illustrated in figures \ref{NearFieldRS} and \ref{FarFieldRS}, which show the scaled radial components of displacement for stations along the El Mayor-Cucapah P axis.  When using eq. \ref{ApproxForward} to fit the entire five years of postseismic displacement we see that the near field displacements, eg. station P501, are accurately predicted, as was also the case when modelling postseismic deformation with an elastic lithosphere.  When looking at displacement in the far field, eg. station P621, we see that eq. \ref{ApproxForward} overestimates the rate of deformation in the later postseismic and underestimates the rate of deformation in the near field.  As we progressively decrease the length of the postseismic timeseries used in the inversion, we see improved fit to the early postseismic period. Due to the low signal to noise ratio for far field stations, it is difficult to determine at what point eq. \ref{ApproxForward} is no longer able to predict the observed displacements without incurring any significant systematic misfit; however, we settle on $t_{\mathrm{end}}=0.8$ years after the earthquake, while acknowledging that the choice is subjective. As noted in \cite{Hines2015}, overestimating $t_{\mathrm{end}}$ will result in a bias towards overestimating $\eta_{\mathrm{eff}}$, while picking a $t_\mathrm{end}$ which is too low will not necessarily result in a biased estimate of $\eta_\mathrm{eff}$; although the uncertainties would be larger. We can then consider our subsequent inferences of $\eta_{\mathrm{eff}}$ to be an upper bound on the lithospheric strength needed to describe the far field rate of deformation during the first 0.8 years of postseismic deformation.

We want to find the discretize parameters describing coseismic slip and the rate of afterslip, $\mathbf{s}$, and viscosity, $\mathbf{\eta_{\mathrm{eff}}}$, that minimize the difference between the each discrete observation of postseismic deformation over the first 0.8 years following the El Mayor-Cucapah earthquake, which is here denoted $\mathbf{u_\mathrm{obs}}$, and the discrete observations predicted in eq. \ref{ApproxForward}, $\mathbf{u_\mathrm{pred}}$. That is, we want to solve

\begin{equation}\label{ObjectiveFunction}
 \mathrm{min}_{\mathbf{s},\mathbf{\eta_\mathrm{eff}}} \left|\left|
 \frac{\mathbf{u_\mathrm{pred}}(s,\eta_\mathrm{eff}) - \mathbf{u_\mathrm{obs}}}
 {\mathbf{\sigma_\mathrm{obs}}}\right|\right|_2^2.
\end{equation} 
Due to inherent nonuniqueness in the solution to eq. \ref{ObjectiveFunction}, we add zeroth order Tikhonov regularization to estimates of slip,
\begin{equation}\label{Misfit}
||\lambda_s \mathbf{s}||_2^2=0,
\end{equation}
and second order Tikhonov regularization to estimates of effective fluidity, 
\begin{equation}
||\lambda_v \nabla \mathbf{\eta_{\mathrm{eff}}}^{-1}||_2^2=0.
\end{equation}
The degree to which we impose the regularization on slip and viscosity is controlled by the penalty parameters $\lambda_s$ and $\lambda_v$.  The penalty parameters are chosen with two L-curve tests, where $\lambda_s$ is first chosen assuming $\lambda_v=0$, and then $\lambda_v$ is determined while keeping $\lambda_s$ at the chosen value.  We note that our goal here is to get a prior constraint on $\eta_{\mathrm{eff}}$ to minimize the amount of searching we have to do in a model space where eq. \ref{GeneralForward} is our forward problem in section \ref{FullInversion}.  Estimates of $\mathbf{s}$ made here will not be used in section \ref{FullInversion}, and so the motivation behind even adding regularization to $\mathbf{s}$ is to ensure that the slip driving viscoelastic relaxation in eq. \ref{ApproxForward} are sensible.

Our inferred estimates of effective viscosities, and corresponding fluidities, are shown in figure \ref{EffectiveViscosity} with their 95\% confidence intervals indicated, which were inferred through bootstrapping. Although fluidities is a sparesly used metric, it is more natural to use fludities rather than viscosity since eq. \ref{InitialRate} is linear with respect to fluidity and so the fluidity represents the amplitude of the viscoelastic signal coming from each discretized layer of the lithosphere.  We note that the magnitude of the uncertainties on viscosity tend to decrease as we increase $\lambda_v$, and our choice our $\lambda_v$ was based on a standard technique used in geophysical inverse problems which has no statistical backing.  It is therefore difficult to interpret the magnitude of uncertainties on viscosity shown in figure \ref{EffectiveViscosity}; although, we do believe that the relative uncertainty between layers is accurately depicted.  The uncertainties in fluidities increases with depth, as would be expected due to decreased sensitivity with depth.  Nevertheless, a robust feature that we see in this that the viscosity below 60 km depth needs to be $\sim 10^{18}$ Pa s to describe the early rate of postseismic deformation for stations greater than $\sim 200$ km from the epicenter. The viscosity of the lower crust is the least well constrained as there is effectively no evidence of relaxation in that layer, meaning that the viscosity is effectively infinite during the first 0.8 years of postseismic, $> 10^{19}$ Pa s. Our inference of a mantle viscosity of $\sim 10^{18}$ Pa s and a relatively stronger lower crust is consistent with the inferred steady state viscosities from \cite{Pollitz2000}, \cite{Pollitz2003}, \cite{Johnson2007}, \cite{Spinler2015}, and \cite{Rollins2015}.

Our initial estimate for coseismic slip and cumulative afterslip over the first 0.8 years after the El Mayor-Cucapah earthquake are shown in figures \ref{InitialCoseismic} and \ref{InitialAfterslip} respectively. To show the how our slip solution is influenced by allowing for a viscoelastic lithopshere, we have also show a coseismic slip and afterslip inferred using the same geometry, discretization, and regularization, but while assuming the lithosphere and ashthenosphere are elastic.  In both coseismic inversion, most of the slip is inferred to be in the Sierra Cucapah and is right lateral with a significant normal component. This is consistent with field studies, \cite{Fletcher2014} as well as coseismic slip models \cite{Wei2011}.  The potency of inferred coseismic slip is $3.32\times 10^{9} \mathrm{m}^3$, which is equivalent to a moment magnitude of 7.28 when assuming a shear modulus of $3.2\times10^{10}$ Pa. The potency of the inferred coseismic slip model when assuming an elastic lithosphere is not significantly different, $3.55\times 10^9 \mathrm{m}^3$.  The striking difference between the two models is that the inferred afterslip on the Indiviso fault segment is significantly larger when not accounting for viscoelasticity.  The potency of afterslip in the elastic model is $2.7\times 10^9 \mathrm{m}^3$, compared to $0.85\times 10^9 \mathrm{m}^3$.  The former is comparable to the size of the mainshock which, as noted by \cite{Rollins2015}, is implausibly large. The significant amount of afterslip on the Indiviso fault inferred by \cite{Rollins2015}, which is predicts broadscale deformation for the GPS stations used in this study, seems to be compensating for unmodeled viscoelastic relaxation at depths of $>60$ km. This inferred transitional depth is consistent with the regionally averaged lithosphere-asthenosphere boundary inferred by \cite{Lekic2011}.  The fact that there is still a substantial amount of deep afterslip inferred on the Indiviso fault even when allowing for viscoelastic relaxation in the lower crust and upper mantle raises the question of whether it is compensating for viscoelastic relaxation that is more localized than we allow for since we only estimate depth dependent variations in viscosity.

\subsection{Full Inversion}\label{FullInversion}

In the previous section we used the efficient inverse method described by \cite{Hines2015} to constrain the effective lithospheric viscosity required to explain the first 0.8 years of postseismic deformation. In this section we use the inferred effective lithospheric viscosities as a prior constraint when searching for rheologic models of the lithosphere which are capable of describing the available 5 years of postseismic data.

Table 1 describes how the inferred effective viscosities from section \ref{InitialInversion} constrain the rheologic parameters for various assumed rheologic models. In this section we perform a series of inversions where we estimate the coseismic slip and afterslip assuming a variety of lithospheric rheologies which are consistent with our findings from \ref{InitialInversion}.  Our forward problem is now eq. \ref{GeneralForward} rather than the approximation given by eq. \ref{ApproxForward}.  As a reference, we compare the best fit solution for each assumed lithospheric model to a the best fit solution assuming an entirely elastic lithosphere. One intuitive criterion for a viscoelastic lithospheric model to be plausible is that it should be able to predict observable deformation better than a lithospheric model which is elastic.

We first assume that the crust and mantle can be described with a Maxwell rheology, and so inferences of $\eta_{\mathrm{eff}}$ are equivalent to inferences of $\eta_{\mathrm{M}}$. We compute $f$ and $g$ from eq. \ref{GeneralForward} using Pylith \cite{Aagaard2009} and assume the same spatial and temporal discretization of $s$ as in section \ref{InitialInversion}. The misfit, defined by eq. \ref{Misfit}, is X which significantly higher than that of the elastic model, X.  It is apparent that most of the misfit comes from the Maxwell model significantly overestimating the rate of deformation after $\sim 3$ years following the earthquake.  Although, the numerous studies have described postseismic deformation in Southern California using a Maxwell visocolatic rheology with viscosities consistent with ours, it is clear that such a model is incapable of describing the entire postseismic time series. As previously noted, \cite{Pollitz2001}, recognized this deficiency in a Maxwell rheology which motivated their exploration of a Burgers rheology upper mantle \cite{Pollitz2003}.
